---
title: "R Notebook"
output: html_notebook
---

```{r}
library(factoextra)
library(corrplot)
library(tidyverse)
library(ggplot2)
library(reshape2)
library(ggcorrplot)
library(openxlsx)
library(ggpubr)
library(MASS)
```


```{r}
data_mall <- read.csv("Mall_Customers.csv")
str(data_mall)
```

```{r}
data_mall <- data_mall[,c("Age","Annual.Income","Spending.Score")]
head(data_mall)
```

```{r}
data_mall_standardize <- scale(data_mall)
apply(data_mall_standardize,2,mean)
apply(data_mall_standardize,2,sd)
```


```{r}
fviz_nbclust(data_mall_standardize,FUNcluster = kmeans,method = "silhouette")
```



```{r}
fviz_nbclust(data_mall_standardize,FUNcluster = kmeans,method = "wss")
```


```{r}
fviz_nbclust(data_mall_standardize,FUNcluster = kmeans, k.max = 20, method = "wss")
```


```{r}
kmeans_mall <- eclust(data_mall, stand = TRUE, FUNcluster = "kmeans", k=4, graph = F)
kmeans_mall$cluster
```

```{r}
kmeans_mall$centers
```

```{r}
aggregate(data_mall,by =list(cluster=kmeans_mall$cluster), FUN = mean)
```


```{r}
fviz_cluster(kmeans_mall)
```

```{r}
pca_mall <- prcomp(data_mall_standardize)
pca_mall$rotation
```
## Hierarchical Clustering Method

```{r}
#complete
fviz_nbclust(data_mall_standardize,FUNcluster = hcut,method = "silhouette",
             hc_method = "complete",hc_metric = "euclidean")
```


```{r}
#average
fviz_nbclust(data_mall_standardize,FUNcluster = hcut,method = "silhouette",
             hc_method = "average",hc_metric = "euclidean")
```


```{r}
#centroid
fviz_nbclust(data_mall_standardize,FUNcluster = hcut,method = "silhouette",
             hc_method = "centroid",hc_metric = "euclidean")
```


```{r}
#ward
fviz_nbclust(data_mall_standardize,FUNcluster = hcut,method = "silhouette",
             hc_method = "ward.D",hc_metric = "euclidean")
```

using dendogram

```{r}
linkage_methods <- c("complete","average","centroid","ward.D")
hc_mall_dend <- lapply(linkage_methods, function(i)
  hclust(dist(data_mall_standardize,method = 'euclidean'),method = i)
  )
#complete
fviz_dend(hc_mall_dend[[1]])
```


```{r}
#average
fviz_dend(hc_mall_dend[[2]])
```

```{r}
#centroid
fviz_dend(hc_mall_dend[[3]])
```

```{r}
#ward
fviz_dend(hc_mall_dend[[4]])
```

Hierarcichal Clustering Implementation

```{r}
hc_mall <- eclust(data_mall,stand = TRUE,FUNcluster = "hclust",k=5,hc_method = "complete",hc_metric = "euclidean",graph = F)
hc_mall$cluster
```

Interpretation of the Formed Cluster

```{r}
aggregate(data_mall,by =list(cluster=hc_mall$cluster),
            FUN = mean)
```

```{r}
fviz_cluster(hc_mall)
```

```{r}
pca_mall <- prcomp(data_mall_standardize)
pca_mall$rotation
```

## Correlation Heatmap

```{r}
corrplot(mpg %>% select(where(is.numeric)) %>% cor,
         method = "pie",type = "lower",
         diag = FALSE)
```

```{r}
mydata <- mtcars[, c(1,3,4,5,6,7)]
head(mydata)
```


```{r}
cormat <- round(cor(mydata),2)
head(cormat)
```


```{r}
melted_cormat <- melt(cormat)
head(melted_cormat)
```


```{r}
ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()
```

```{r}
# Get lower triangle of the correlation matrix
get_lower_tri<-function(cormat){
  cormat[upper.tri(cormat)] <- NA
  return(cormat)
}
# Get upper triangle of the correlation matrix
get_upper_tri <- function(cormat){
  cormat[lower.tri(cormat)]<- NA
  return(cormat)
}
upper_tri <- get_upper_tri(cormat)
upper_tri
```

```{r}
melted_cormat <- melt(upper_tri, na.rm = TRUE)
# Heatmap
ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()
```

```{r}
#helper function to reorder correlation matrix:
reorder_cormat <- function(cormat){
# Use correlation between variables as distance
dd <- as.dist((1-cormat)/2)
hc <- hclust(dd)
cormat <-cormat[hc$order, hc$order]
}
```


```{r}
# Reorder the correlation matrix
cormat <- reorder_cormat(cormat)
upper_tri <- get_upper_tri(cormat)
# Melt the correlation matrix
melted_cormat <- melt(upper_tri, na.rm = TRUE)
# Create a ggheatmap
ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
    name="Pearson\nCorrelation") +
  theme_minimal()+ # minimal theme
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()
# Print the heatmap
print(ggheatmap)
```

```{r}
ggheatmap + 
geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))
```

## Dimension Reduction

```{r}
data_women_records <- read.xlsx("women_track_records.xlsx")
head(data_women_records)
```

```{r}
rownames(data_women_records) <- data_women_records$country
data_women_records <- data_women_records[,-8]
```

```{r}
cor_women <- cor(data_women_records)
cor_women
```

```{r}
ggcorrplot(cor_women,type="lower",lab = TRUE)
```

Matriks korelasi tidak hanya melihat rataan tetapi juga kekuatan hubungan antar peubahnya 
Matriks kovarian tidak melihat seberapa kuat antar hubungan tetapi melihat arahnya saja, positif atau negatif


```{r}
pca_women_records <- prcomp(data_women_records,scale.=TRUE,center=TRUE)
summary(pca_women_records)
```

83% PC1 bisa menjelaskan variable kita. dua komponen menjelaskan lebih banyak dibandingkan yang lain. 

```{r}
fviz_screeplot(pca_women_records,geom="line")
```

```{r}
pca_women_records$rotation
```

```{r}
pca_women_records$x
```

```{r}
fviz_pca_ind(pca_women_records,col.ind = "darkred")
```

```{r}
data(swiss)
head(swiss)
```


```{r}
# calculate distance matrix
distance_swiss <- dist(swiss, method = "euclidean")
# Metric MDS
metricMDS_swiss <- cmdscale(d = distance_swiss,k = 2)
colnames(metricMDS_swiss) <- c("Dim.1","Dim.2")
metricMDS_swiss <- as.data.frame(metricMDS_swiss)
head(metricMDS_swiss)
```


```{r}
KruskalMDS_swiss <- isoMDS(d = distance_swiss,k = 2)
```

```{r}
KruskalMDS_swiss <- as.data.frame(KruskalMDS_swiss$points)
colnames(KruskalMDS_swiss) <- c("Dim.1","Dim.2")
head(KruskalMDS_swiss)
```

```{r}
ggscatter(metricMDS_swiss,x="Dim.1",y="Dim.2",color = "steelblue",
          label = rownames(swiss),
          repel = TRUE)
```

```{r}
ggscatter(KruskalMDS_swiss,x="Dim.1",y="Dim.2",color = "steelblue",
          label = rownames(swiss),
          repel = TRUE)
```



