---
title: "Tugas Praktikum STA1563: Eksplorasi dan Visualisasi Data"
author: "Alfa Nugraha^[Mahasiswa Pascasarjana IPB, alfanugraha@apps.ipb.ac.id]"
date: ""
output:
  html_notebook:
    toc_float: true
    df_print: paged
    toc: true
    theme: flatly
    highlight: zenburn
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(skimr)
library(ggplot2)
library(dplyr)
library(DataExplorer)
library(reshape2)
```


## Dataset INDO-DAPOER

The Indonesia Database for Policy and Economic Research (INDO-DAPOER) contains relevant economic and social indicators at the province- and district-level, which span across four main categories: fiscal, economic, social and demographic, as well as infrastructure. The data was accessed through this link: https://databank.worldbank.org/reports.aspx?source=1266

The column of dataset contains:

- `ProvName`: Province name
- `ProvCode`: Province code
- `SeriesName`: Indicator name of fiscal, economic, social and demographic, and infrastructure
- `SeriesCode`: Indicator code
- `YR1976` - `YR2020`: The value of each indicator from 1976-2020


### Data preparation

```{r}
indo_dapoer <- read.csv2("indo_dapoer.csv", stringsAsFactors = T)
indo_dapoer
```

### Summary

```{r}
str(indo_dapoer)
```

```{r}
skim(indo_dapoer)
```

```{r}
plot_intro(indo_dapoer)
```


## Exploration

### Villages with road

- Source: BADAN PUSAT STATISTIK - Statistics Indonesia, Village Census (PODES)
- Topic: Infrastructure Indicators
- Unit of Measures: Percent
- Periodicity: Triennial
- General Comments: Data only available once in three years: 1996, 2000, 2003, 2005, 2008, 2011, 2014, 2018.
- Other Notes: World Bank Staff Calculation based on Village Census (PODES) - Statistics Indonesia
- Type of road: Asphalt, Dirt, Gravel, Other


```{r}
rod_vilg <- indo_dapoer %>% 
  filter(SeriesCode %in% c("ROD.VILG.ASPH.ZS", "ROD.VILG.DIRT.ZS", "ROD.VILG.GRAVL.ZS", "ROD.VILG.OTHR.ZS")) %>%
  select(c(ProvCode, SeriesCode, YR1996, YR2000, YR2003, YR2005, YR2008, YR2011, YR2014, YR2018)) %>%
  mutate(Total = apply(rod_vilg[3:8], 1, sum))
rod_vilg
```

```{r}
rod_vilg_cast <- dcast(rod_vilg, ProvCode ~ SeriesCode, value.var = 'Total')
rod_vilg_sum <- data.frame(road=c("Asphalt", "Dirt", "Gravel", "Other"), total=apply(rod_vilg_cast[, 2:5], 2, sum))
ggplot(rod_vilg_sum, aes(road, total)) + geom_col()
```

```{r}
rod_vilg_y_cast <- dcast(rod_vilg, ProvCode ~ SeriesCode)
rod_vilg_y_cast
# apply(rod_vilg_cast[, 2:5], 2, sum)
# ggplot(rod_vilg[1,], aes(x=ProvCode, y=Total)) + geom_point() + geom_smooth()
```



```{r}
dense <- density(rod_vilg$Total, bw=1, kernel="epanechnikov")
hist(rod_vilg$Total, freq = FALSE, breaks = 15, col = "skyblue1", main = "", xlab = "Total")
lines(dense, col="blue", lwd=2, main="", ylim=c(0, 0.09))
```


```{r}
ggplot(rod_vilg, aes(x=Total, color=SeriesCode, fill=SeriesCode)) +
  geom_bar() + 
  scale_x_binned()
```
```{r}
ggplot(rod_vilg, aes(x=Total, after_stat(density), colour=SeriesCode)) +
  geom_freqpoly(binwidth = 500)
```


```{r}
boxplot(rod_vilg$Total ~ rod_vilg$ProvCode,
        col= c("coral","green","skyblue", "red"),
        ylab= "Mother Age (year)",
        xlab = "Race")
```


#### Asphalt (in % of total villages)

```{r}
rod_vilg_asph <- rod_vilg %>% 
  filter(SeriesCode %in% c("ROD.VILG.ASPH.ZS")) 
rod_vilg_asph
```

```{r}
ggplot(rod_vilg, aes(x=ProvCode, y=Total)) + geom_point() + geom_smooth()
```

```{r}
hist(rod_vilg_asph$Total)
```


```{r}
boxplot(rod_vilg_asph$Total, horizontal = T, col = 'coral', xlab="Asphalt (in % of total villages)")
```

```{r}
boxplot(rod_vilg_asph$YR1996, horizontal = T, col = 'coral', xlab="Asphalt (in % of total villages)")
```

## References

- Data Catalog. https://datacatalog.worldbank.org/search/dataset/0041056
